En garde! Unguarded iteration for reversible computation in the Delay monad\thanks{Niccol\`o Veltri was supported by a research grant (13156) from VILLUM FONDEN.}

### 2019-04-30 mpc19/paper.tex
Added a conclusion!

\section{TODO} % (fold)
\label{sec:todo}
Robin:
\begin{itemize}
  \item Conclusion
  \begin{itemize}
    \item Related work: Kleisli category of the Delay monad is a join
    restriction category, inverse category of a join restriction category is a
    join inverse category; already known to be dagger-traced under certain
    conditions.
    \item Future work: Semantics of time invertible programming. Extension of
    the work on the Kleisli category of the Delay monad beyond $\Set$ to other
    (more general) categories, e.g. topoi.
  \end{itemize}
\end{itemize}

% section todo (end)

@@

@@

\section{Conclusions}

>>>>

@@

By abuse of notation, we use here $\cdot \times \cdot$ (respectively $\cdot +
\cdot$) to refer to the product (respectively coproduct) in $\Dapprox$ even
though it fails to be a product (respectively coproduct) in $\Inv\Dapprox$.
However, both of these are symmetric monoidal products in $\Inv\Dapprox$, so
their use as objects of $\Inv\Dapprox$ in the interpretation above is justified.

@@

\section{Conclusions}
In this paper, we have extended the work of Carette and Sabry~\cite{CaretteS16}
to a (fully formalized) two-level calculus of \Pio{} programs and program
equivalences. Key in this effort was the use of the Kleisli category of the
$\Delay$ monad on $\Set$ under weak bisimilarity, which turned out to support
iteration via a trace that preserves all partial isomorphisms, in this way
giving semantics to the dagger trace of \Pio. Further, the work was formalized
using Agda 2.6.0.

The notion of computable partiality offered the $\Delay$ monad and its
restriction category structure is well
studied~\cite{VeltriPhD,UustaluV17,ChapmanUV19}. Indeed, though it could be
established from previous work (see also \cite{KaarsgaardAG17}) that this
category was traced, it was unclear whether the trace preserved partial
isomorphisms -- a crucial point in its applicability to interpreting \Pio. The
presented solution further avoids CPO-enrichment (as otherwise used in both
\cite{UustaluV17} and \cite{KaarsgaardAG17}), and its issues in relation to
formalization, in favor of a more direct definition of the trace from the Elgot
iterator.

Though the Kleisli category of the $\Delay$ monad on $\Set$ is well studied,
comparatively less is known about this monad on other categories. It could be
interesting to study under which conditions its iterator exist -- e.g., whether
this is still the case when $\Set$ is replaced with an arbitrary topos. Another
avenue concerns the study of \emph{time invertible} programming languages:
Though not immediately clear in the current presentation, the trace on
$\Inv\Dapprox$ is not just reversible (in the sense that it preserves partial
isomorphisms) but in fact time invertible, in the sense that the number of
computation steps needed to perform $\trace (\dagger f)$ is \emph{precisely}
the same as what is needed to perform $\trace~f$ on any input. Since the
$\Delay$ monad conveniently allows the counting of computation steps, we
conjecture that this is an ideal setting in which to study such intentional
semantics of reversible programming languages.
